Discrete Quantum Theories and Computing

### 2019-01-13 dissertation.tex
dissertation.lyx Preface, 5.3.1 Finite-Precision Extension of Gleason's Theorem, and Chapter 6 Further Discussion Updating...

\begin{preface*}
\begin{itemize}
\item The latest PDF version of this thesis is in GitHub: \code{https://git.io/fxbuG}.
\item Its source code is typeset using \LyX ~\cite{LyX} in \code{https://git.io/fxbuE}
and the exported \LaTeX{} file is in \code{https://git.io/fhZBW}.
\item Any comments can be left in the Issue part of the GitHub repository:
\code{https://git.io/fxbug}.
\end{itemize}
Following explains how to compile the PDF from the \LaTeX{} source:
\begin{enumerate}
\item \label{enu:Clone-the-git}Clone the git repository of this thesis
\code{https://git.io/fhZP5} into your local directory and the thesis
document class \code{https://github.com/yuttai/iuphd} in a location
where the system can find.
\item In the local directory of this thesis, run \code{xelatex dissertation.tex}.
\item Run \code{biber dissertation}.
\item Run \code{xelatex dissertation.tex} several times until all the references
are resolved.
\end{enumerate}
The previous steps only tested in Microsoft Windows. Please report
me if there is any difficulty to compile it. Besides, to compile a
PDF file from the \LyX{} source, just finish step~\ref{enu:Clone-the-git}
and click the toolbar button \includegraphics[width=1em]{buffer-view}.\end{preface*}

@@

To that end, we introduce the concept of \emph{interval maps}, which
we can use to construct a consistent family of QIVPMs. An interval
map $f:\left[0,1\right]\rightarrow\mathscr{I}$ maps every real-valued
probability $x\in\left[0,1\right]$ to a set of intervals $f\left(x\right)=\left[\ell,r\right]$
containing $x$, where $\left[0,1\right]$ denotes the set of real-valued
probabilities (this should not be confused with the interval-valued
probability $\unknown$). We also need a notion of \emph{norm} to
quantify the distance between (pure or mixed) states. The norm of
a pure state $\rho=\proj{\psi}$ is defined as usual by $\left\Vert \psi\right\Vert =\sqrt{\ip{\psi}{\psi}}$.
For any given Hermitian operator~$A$, we choose the operator norm
$\left\Vert A\right\Vert =\max_{\left\Vert \psi\right\Vert =1}\left\Vert A\ket{\psi}\right\Vert $,
which is also known as the $2$-norm or the spectral norm~\cite{RobertsVarberg1973,peres1995quantum,GolubVanLoan1996,Foucart2012}.
In fact, for any such matrix, including the density matrix~$\rho$,
this norm is the maximum absolute value of its eigenvalues. Then,
a finite-precision extension of Gleason's theorem can be stated as
follows.

\begin{thm}[Finite-Precision Extension of the Gleason Theorem]\label{thm:Finite-precision-Gleason}Let
$f:\left[0,1\right]\rightarrow\mathscr{I}$ be an interval map and
let the composition $f\circ\muB_{\rho}$ be a QIVPM, where $\muB_{\rho}$
is the probability measure induced by the Born rule for a given state~$\rho$.
Let $\alpha$ be the maximum length of intervals in $\mathscr{I}$.
If a state $\rho'$ is consistent with $f\circ\muB_{\rho}$ on all
events, i.e., $\rho'\in\coreBorn\left(f\circ\muB_{\rho},\events\right)$,
then the norm of their difference is bounded by $\alpha$, i.e., $\left\Vert \rho-\rho'\right\Vert \le\alpha$.\end{thm}

\begin{proof}Given a state~$\rho'$ consistent with $f\circ\muB_{\rho}$,
we have $\muB_{\rho'}\left(\proj{\psi}\right)\in f\left(\muB_{\rho}\left(\proj{\psi}\right)\right)$
for any one-dimensional projector $P=\proj{\psi}$. Since the maximum
length of the intervals in $\mathscr{I}$ is $\alpha$, it is also
the upper bound of the difference: 
\begin{equation}
\left|\muB_{\rho'}\left(\proj{\psi}\right)-\muB_{\rho}\left(\proj{\psi}\right)\right|=\left|\melem{\psi}{\rho-\rho'}{\psi}\right|\le\alpha\,.
\end{equation}
Since $\rho-\rho'$ is Hermitian, $\max_{\left\Vert \psi\right\Vert =1}\left|\melem{\psi}{\rho-\rho'}{\psi}\right|$
is the maximum absolute value of the eigenvalues of $\rho-\rho'$~\cite{544199},
and equal to $\left\Vert \rho-\rho'\right\Vert $~\cite{GolubVanLoan1996,Foucart2012}.
Hence, $\left\Vert \rho-\rho'\right\Vert \le\alpha$.\end{proof}

@@

\begin{comment}
When people proved the original Gleason theorem, people usually exploited
the geometrical structure of real 3-dimensional Hilbert space \cite{gleason1957,peres1995quantum,RichmanBridges1999,Hamhalter2013}.
Since our finite-precision extension of the Gleason theorem only applies
on a class of QIVPMs, we might want to ask how to modify these geometrical
arguments to have a Gleason-type theorem for general QIVPMs. We will
further study the tensor product structure among QIVPMs which is essential
for defining product and entangled states, and serves the basis to
discuss quantum nonlocality~\cite{Bell1964,Redhead1987-REDINA,peres1995quantum,Jaeger2007}
and quantum computing with QIVPMs. Finally, we want to improve the
discrete quantum theories to consider QIVPMs over finite fields in
future research.
\end{comment}
As we discussed in the previous section, the finite-precision extension
of the Gleason theorem, Thm.~\ref{thm:Finite-precision-Gleason},
only applies on the QIVPMs composed by interval maps $f:\left[0,1\right]\rightarrow\mathscr{I}$
and quantum probability measures $\muB_{\rho}:\events\rightarrow\left[0,1\right]$:
the states consistent with the composition $f\circ\muB_{\rho}$ converges
to $\rho$ as the maximum length of intervals in $\mathscr{I}$ shrinking
to $0$. In contrast, the wildest QIVPM defined in Table~\ref{tab:non-Born-QIVPM}
maps to $\mathscr{I}_{0}$ whose intervals' maximum length is $1$,
the longest possible length. This provides another possibility that
shirking the maximum length of intervals itself could result the QIVPMs
mapping to these intervals having non-empty cores. More precisely,
we have the following question.

\begin{question}\label{question:approximation-Gleason}Given a Hilbert
space $\Hilb$ of dimension $D\geq3$, is there an $\varepsilon>0$
such that for all QIVPM $\bar{\mu}:\events\rightarrow\mathscr{I}$
satisfying the maximum length of intervals in $\mathscr{I}$ less
than $\varepsilon$, $\bar{\mu}$ must have an non-empty core, i.e.,
$\coreBorn\left(\bar{\mu},\events\right)\ne\emptyset$.\end{question}

@@

To better understand this question, consider the $D=3$ situation
which is largely simplified by the fact that any two-dimensional projectors
can be expressed as the complement of a one-dimensional projector,
and by Eq.~(\ref{eq:QIVPM-complement}) so does their interval-valued
probabilities, i.e., $\bar{\mu}\left(\mathbb{1}-\proj{\psi}\right)=\necess-\bar{\mu}\left(\proj{\psi}\right)$.

@@

we can decompose interval-valued probabilities on these states into
its left-end and right-end, where $\left[\frameL\left(\ket{\psi}\right),\frameR\left(\ket{\psi}\right)\right]=\bar{\mu}\left(\proj{\psi}\right)$.

@@

Since the maximum length of intervals in $\mathscr{I}$ is less than
$\varepsilon$, $\left[\frameL\left(\ket{\psi}\right),\frameR\left(\ket{\psi}\right)\right]\in\mathscr{I}$
implies $\frameR\left(\ket{\psi}\right)-\frameL\left(\ket{\psi}\right)<\varepsilon$.
Then, the largest term in the previous inequality $\frameR\left(\ket{\psi_{1}}\right)+\frameR\left(\ket{\psi_{2}}\right)$

@@

for any $D$-dimensional projector~$P$, where $\ket{\Phi_{j}}\in\mathcal{H}$

@@

ask what $\frameL$ would look like with a positive $\varepsilon$.

With different settings, whether there is an approximate version of
Gleason's theorem in general was asked by Sam Sanders in constructivenews~\cite{Sanders2013},
and there is no clear answer for his question as far as I understand.
To have an idea how hard this question could be, recall in Sec.~\ref{sec:fuzzy}
we state that a quantum probability space is glued by a family of
classical probability spaces. This is similar to the situation that
a manifold is glued by many local coordinates. When the curvature
of each small piece is ``exactly'' the same and positive, Killing-Hopf
theorem asserts this manifold is a sphere, but little can we say even
if the curvature has only small deviation from constant. Similar situation
might happen when approximating Gleason's theorem. Since the sum of
$\frameL$ on each local classical probability space is ``exactly''
the same and equal to $1$, Gleason's theorem can assert $\frameL$
can be expressed as Eq.~\ref{eq:Gleason-for-CP2}. However, when
each local classical probability space becomes imprecise, a general
$\frameL$ might be as wild as we can image, and we might say little
about its global property.

>>>>

\begin{preface*}
\begin{itemize}
\item The latest PDF version of this thesis is in GitHub: \code{https://git.io/fxbuG}.
\item Its source code is typeset using \LyX ~\cite{LyX} in \code{https://git.io/fxbuE}
and the exported \LaTeX{} file is in \code{https://git.io/fhZBW}.
\item Any comments can be left in the Issue part of the GitHub repository:
\code{https://git.io/fxbug}.
\end{itemize}
Following explains how to compile the PDF from the \LaTeX{} source:
\begin{enumerate}
\item \label{enu:Clone-the-git}Clone the git repository of this thesis
\code{https://git.io/fhZP5} into your local directory and the thesis
document class \code{https://github.com/yuttai/iuphd} in a location
where the system can find.
\item In the local directory of this thesis, run \code{xelatex dissertation.tex}.
\item Run \code{biber dissertation}.
\item Run \code{xelatex dissertation.tex} several times until all the references
are resolved.
\end{enumerate}
The previous steps only tested in Microsoft Windows with MiK\TeX ~2.9.
Please report me if there is any difficulty to compile it. Besides,
to compile a PDF file from the \LyX{} source, just finish step~\ref{enu:Clone-the-git},
open \code{dissertation.lyx} in \LyX , and click the toolbar button
\includegraphics[width=1em]{buffer-view}.\end{preface*}

@@

To that end, we introduce the concept of \emph{interval maps}, which
we can use to construct a consistent family of QIVPMs. An interval
map $f:\left[0,1\right]\rightarrow\mathscr{I}$ maps every real-valued
probability $x\in\left[0,1\right]$ to a set of intervals $f\left(x\right)=\left[\ell,r\right]$
containing $x$, where $\left[0,1\right]$ denotes the set of real-valued
probabilities (this should not be confused with the interval-valued
probability $\unknown$). Let $\left|\mathscr{I}\right|$ be the maximum
length of intervals in $\mathscr{I}$. We also need a notion of \emph{norm}
to quantify the distance between (pure or mixed) states. The norm
of a pure state $\rho=\proj{\psi}$ is defined as usual by $\left\Vert \psi\right\Vert =\sqrt{\ip{\psi}{\psi}}$.
For any given Hermitian operator~$A$, we choose the operator norm
$\left\Vert A\right\Vert =\max_{\left\Vert \psi\right\Vert =1}\left\Vert A\ket{\psi}\right\Vert $,
which is also known as the $2$-norm or the spectral norm~\cite{RobertsVarberg1973,peres1995quantum,GolubVanLoan1996,Foucart2012}.
In fact, for any such matrix, including the density matrix~$\rho$,
this norm is the maximum absolute value of its eigenvalues. Then,
a finite-precision extension of Gleason's theorem can be stated as
follows.

\begin{thm}[Finite-Precision Extension of the Gleason Theorem]\label{thm:Finite-precision-Gleason}Let
$f:\left[0,1\right]\rightarrow\mathscr{I}$ be an interval map and
let the composition $f\circ\muB_{\rho}$ be a QIVPM, where $\muB_{\rho}$
is the probability measure induced by the Born rule for a given state~$\rho$.
If a state $\rho'$ is consistent with $f\circ\muB_{\rho}$ on all
events, i.e., $\rho'\in\coreBorn\left(f\circ\muB_{\rho},\events\right)$,
then the norm of their difference is bounded by $\left|\mathscr{I}\right|$,
i.e., $\left\Vert \rho-\rho'\right\Vert \le\left|\mathscr{I}\right|$.\end{thm}

\begin{proof}Given a state~$\rho'$ consistent with $f\circ\muB_{\rho}$,
we have $\muB_{\rho'}\left(\proj{\psi}\right)\in f\left(\muB_{\rho}\left(\proj{\psi}\right)\right)$
for any one-dimensional projector $P=\proj{\psi}$. Since the maximum
length of the intervals in $\mathscr{I}$ is $\left|\mathscr{I}\right|$,
it is also the upper bound of the difference: 
\begin{equation}
\left|\muB_{\rho'}\left(\proj{\psi}\right)-\muB_{\rho}\left(\proj{\psi}\right)\right|=\left|\melem{\psi}{\rho-\rho'}{\psi}\right|\le\left|\mathscr{I}\right|\,.
\end{equation}
Since $\rho-\rho'$ is Hermitian, $\max_{\left\Vert \psi\right\Vert =1}\left|\melem{\psi}{\rho-\rho'}{\psi}\right|$
is the maximum absolute value of the eigenvalues of $\rho-\rho'$~\cite{544199},
and equal to $\left\Vert \rho-\rho'\right\Vert $~\cite{GolubVanLoan1996,Foucart2012}.
Hence, $\left\Vert \rho-\rho'\right\Vert \le\left|\mathscr{I}\right|$.\end{proof}

@@

\section{Gleason's Theorem for General QIVPMs}

\begin{comment}
When people proved the original Gleason theorem, people usually exploited
the geometrical structure of real 3-dimensional Hilbert space \cite{gleason1957,peres1995quantum,RichmanBridges1999,Hamhalter2013}.
Since our finite-precision extension of the Gleason theorem only applies
on a class of QIVPMs, we might want to ask how to modify these geometrical
arguments to have a Gleason-type theorem for general QIVPMs. We will
further study the tensor product structure among QIVPMs which is essential
for defining product and entangled states, and serves the basis to
discuss quantum nonlocality~\cite{Bell1964,Redhead1987-REDINA,peres1995quantum,Jaeger2007}
and quantum computing with QIVPMs. Finally, we want to improve the
discrete quantum theories to consider QIVPMs over finite fields in
future research.
\end{comment}
Recall in previous section, we denote the maximum length of intervals
in $\mathscr{I}$ as $\left|\mathscr{I}\right|$. The finite-precision
extension of the Gleason theorem, Thm.~\ref{thm:Finite-precision-Gleason},
only applies on the QIVPMs of the form $f\circ\muB_{\rho}$ and states
that as $\left|\mathscr{I}\right|$ shrinks to $0$, the composition
$f\circ\muB_{\rho}$ would be better in the sense that the states
consistent with the composition $f\circ\muB_{\rho}$ converges to
$\rho$. In contrast, the wildest QIVPM defined in Table~\ref{tab:non-Born-QIVPM}
has the range~$\mathscr{I}_{0}$ with $\left|\mathscr{I}_{0}\right|=1$,
the longest possible length. This suggests that shrinking the length
$\left|\mathscr{I}\right|$ might regularize QIVPMs mapping to $\mathscr{I}$
as the following question.

\begin{question}\label{question:approximation-Gleason}Given a Hilbert
space $\Hilb$ of dimension $D\geq3$, is there an $\varepsilon>0$
such that for all QIVPM $\bar{\mu}:\events\rightarrow\mathscr{I}$
satisfying $\left|\mathscr{I}\right|<\varepsilon$, $\bar{\mu}$ must
have an non-empty core, i.e., $\coreBorn\left(\bar{\mu},\events\right)\ne\emptyset$.\end{question}

@@

To better understand this question, consider the $D=3$ situation,
where any two-dimensional projectors can be expressed as the complement
of a one-dimensional projector, and by Eq.~(\ref{eq:QIVPM-complement})
so does their interval-valued probabilities, i.e., $\bar{\mu}\left(\mathbb{1}-\proj{\psi}\right)=\necess-\bar{\mu}\left(\proj{\psi}\right)$.

@@

we can decompose the interval-valued probabilities on its projectors
$\proj{\psi}$ into the left-end and right-end on $\ket{\psi}$, i.e.,
$\bar{\mu}\left(\proj{\psi}\right)=\left[\frameL\left(\ket{\psi}\right),\frameR\left(\ket{\psi}\right)\right]$.

@@

Since $\left|\mathscr{I}\right|<\varepsilon$, the length of every
interval $\left[\frameL\left(\ket{\psi}\right),\frameR\left(\ket{\psi}\right)\right]$
is bounded by $\varepsilon$ as well, which implies the largest term
in the previous inequality $\frameR\left(\ket{\psi_{1}}\right)+\frameR\left(\ket{\psi_{2}}\right)$

@@

for any state $\ket{\psi}\in\CP{2}$, where $\ket{\Phi_{j}}\in\mathcal{H}$

@@

then ask what $\frameL$ would look like with a positive $\varepsilon$.

With different settings, whether there is an approximate version of
Gleason's theorem in general was asked by Sam Sanders in constructivenews~\cite{Sanders2013},
and there is no clear answer for his question as far as I understand.
To have an idea how hard this question could be, recall in Sec.~\ref{sec:fuzzy}
we state that a quantum probability space is glued by a family of
classical probability spaces. This is similar to the situation that
a manifold is glued by many local coordinates. When each small piece
has \emph{exactly} the same and positive curvature, the Killing-Hopf
theorem asserts this manifold is a sphere, but little can we say even
if the curvature has only small deviation from constant. Similar situation
might happen when approximating Gleason's theorem. Since the sum of
$\frameL$ on each ``local'' classical probability space, $\sum_{i=0}^{2}\frameL\left(\ket{\psi_{i}}\right)$,
is \emph{exactly} the same and equal to $1$, Gleason's theorem can
assert $\frameL$ can be expressed as Eq.~(\ref{eq:Gleason-for-CP2}).
However, when each local classical probability space becomes imprecise,
a general $\frameL$ might be as wild as we can imagine, and we might
say little about its global property.

\section{And Beyond\ldots}
